Adaptive evolution with aneuploidy and mutation

### 2022-03-14 ms/ms.tex
WAIC update

\paragraph{Model comparison.} The main criterion that we use for model selection is WAIC, the widely applicable information criterion ~\citep{gelman2013bayesian},
\begin{equation} \label{eq:WAIC}
\begin{aligned}
\mathit{WAIC}(\theta) &\ =\ 
-2\log\mathbb{E}[\likelihood(\theta)]\ +\ 2\mathbb{V}[\log\likelihood(\theta)]
\end{aligned}
\end{equation}
where $\theta$ is a parameter vector, and $\mathbb{E}[\cdot]$ and $\mathbb{V}[\cdot]$ are the expectation and variance taken over the posterior distribution, which in practice are approximated using 50,000 samples from the posterior KDE. %TODO IK WAIC calculation is still under consideration, update it here after
WAIC values are scaled as a deviance measure: lower values imply higher predictive accuracy~\citep{Kass1995}.

@@

Compared to other priors we tested, this prior produced lower WAIC, better posterior prediction plots, and more stable parameter estimates, as follows.
First, we tried to use additional growth curves of \eumt\; (\emph{refined} strain from \citet{Yona2012}) and \anwt\; in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$ ~(\Cref{fig:growth-curves}L). The same prior was used for $w_{2n*}/w_{2n+1*}$. This prior resulted in WAIC 71.91, compared to 67.87 with the above informative prior. % generated with fixed-mutation-rate-and-WAICs.ipynb TODO IK update the WAIC with 1000 reps run
We also tested an uninformative uniform prior with $\mathit{U}(1,6)$, for (i) all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$, or (ii) only for $w_{2n+1^*}$, $w_{2n^*}$, using the above informative prior for $w_{2n+1}$. In both cases the inference algorithm failed to converge.

@@

%TODO YR read this 
%% generated with fixed-mutation-rate.ipynb 
% TODO IK run for all updated files
\begin{table}[h]
\centering
\caption{
\textbf{WAIC values for various model specifications.}}
\pgfplotstabletypeset[
    col sep=semicolon,
    string type,
    every head row/.style={before row=\hline,after row=\hline},
    every last row/.style={after row=\hline},
    ]{../figures/Table_WAIC.csv} 
\caption*{
	TODO caption
	%TODO IK caption; what is tau in models 1-5?
} 
\label{table:WAIC}
\end{table}

>>>>

\paragraph{Model comparison.} The main criterion that we use for model selection is WAIC, the widely applicable information criterion ~\citep{gelman2013bayesian},
\begin{equation} \label{eq:WAIC}
\begin{aligned}
\mathit{WAIC}(\theta) &\ =\ 
-2\log\mathbb{E}[\likelihood(\theta)]\ +\ 2\mathbb{V}[\log\likelihood(\theta)]
\end{aligned}
\end{equation}
where $\theta$ is a parameter vector, and $\mathbb{E}[\cdot]$ and $\mathbb{V}[\cdot]$ are the expectation and variance taken over the posterior distribution, which in practice are approximated using 50,000 samples from the posterior KDE.
WAIC values are scaled as a deviance measure: lower values imply higher predictive accuracy~\citep{Kass1995}.

@@

Compared to other priors we tested, this prior produced lower WAIC, better posterior prediction plots, and more stable parameter estimates, as follows.
First, we tried to use additional growth curves of \eumt\; (\emph{refined} strain from \citet{Yona2012}) and \anwt\; in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$ ~(\Cref{fig:growth-curves}L). The same prior was used for $w_{2n*}/w_{2n+1*}$. This prior resulted in WAIC -57, compared to 295 with the above informative prior. % generated with WAICs.ipynb TODO IK update the WAIC with 1000 reps run for the new prior
%TODO IK: rewrite text about alternative prior, it has lower WAIC, different estimates for mutation rate etc...maybe it should be moved to results section
We also tested an uninformative uniform prior with $\mathit{U}(1,6)$, for (i) all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$, or (ii) only for $w_{2n+1^*}$, $w_{2n^*}$, using the above informative prior for $w_{2n+1}$. In both cases the inference algorithm failed to converge.

@@

%TODO YR read this 
%% generated with WAIC.ipynb 
\begin{table}[h]
\centering
\caption{
\textbf{WAIC values for various model specifications.}}
\pgfplotstabletypeset[
    col sep=semicolon,
    string type,
    every head row/.style={before row=\hline,after row=\hline},
    every last row/.style={after row=\hline},
    ]{../figures/Table_WAIC.csv} 
\caption*{
	TODO caption
	%TODO IK caption; what is tau in models 1-5?
} 
\label{table:WAIC}
\end{table}
